---
title: "Final Exam"
author: "Mastoras Giannis"
date: "2023-01-30"
output: html_document
---

```{r,include=FALSE}
library(readr)
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(sf)
library(leaflet)
library(cshapes)
library(htmltools)
```

AFTER INCLUDING THE NECESSARY LIBRARIES, WE LOAD OUR DATA AND SPLIT IT'S 
COLUMNS TO CREATE A DATAFRAME.

```{r}
data <- read_csv("NYPD_Shooting_Incident_Data__Historic_.csv")

# First we make our data more visible and check for NA's
data <- data %>%
  separate(`INCIDENT_KEY;OCCUR_DATE;OCCUR_TIME;BORO;PRECINCT;JURISDICTION_CODE;LOCATION_DESC;STATISTICAL_MURDER_FLAG;PERP_AGE_GROUP;PERP_SEX;PERP_RACE;VIC_AGE_GROUP;VIC_SEX;VIC_RACE;X_COORD_CD;Y_COORD_CD;Latitude;Longitude;Lon_Lat`,
          sep = ";",into = c("INCIDENT_KEY","OCCUR_DATE","OCCUR_TIME","BORO",
                    "PRECINCT","JURISDICTION_CODE","LOCATION_DESC",
                    "STATISTICAL_MURDER_FLAG","PERP_AGE_GROUP","PERP_SEX",
                    "PERP_RACE","VIC_AGE_GROUP","VIC_SEX","VIC_RACE","X_COORD_CD",
                    "Y_COORD_CD","Latitude","Longitude","Lon_Lat"))
sum(is.na(data))
```

WE SEE THAT WE HAVE 0 NA's. HOWEVER WE CAN SEE THAT THERE ARE SOME MISSING 
VALUES, WHICH WE WILL MANIPULATE IF IT'S NEEDED IN THE PLOTS.

PLOT 1

IN PLOT 1 WE SEE A BARPLOT THAT INDICATES US WITH THE TOTAL NUMBER OF SHOOTINGS
IN EVERY SINGLE REGION OF NEW YORK.

```{r}
plot1 <- data %>%
  group_by(BORO) %>%
  summarise(Total_incidents_in_region = n()) %>%
  ggplot(aes(x=BORO,y=Total_incidents_in_region))+
  geom_bar(aes(x=BORO),stat = "identity",color="black",fill="grey" ) +
  labs(title = 'Total Events in every Region') +
  scale_y_continuous(breaks=seq(0,12000,1000))
plot1
```

PLOT 2

IN PLOT 2 WE SEPERATE THE ABOVE BARPLOT AND DISTINGUISH EVERY VICTIM'S RACE. WE
SEE THAT MOST OF THE VICTIMS WERE BLACK.

```{r}
plot2 <- data %>%
  group_by(BORO,VIC_RACE) %>%
  summarise(Total_incidents_in_region = n()) %>%
  ggplot(aes(x=BORO,y=Total_incidents_in_region))+
  geom_bar(aes(x=BORO,fill=VIC_RACE),stat = "identity") +
  labs(title = 'Total Events in every Region and Victim Race') +
  scale_y_continuous(breaks=seq(0,12000,1000))
plot2
```

PLOT 3

IN PLOT 3 WE TRY TO SEE IF THERE IS A CORRELATION BETWEEN VICTIM'S AND SHOOTER'S
RACE.

```{r}
plot3<-data %>%
  ggplot() +
  geom_count(mapping = aes(x = PERP_RACE, y = VIC_RACE)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title = 'Correlation Between Victim and Perp Race')
plot3
```

PLOT 4

IN PLOT 4 WE HAVE A NEW CORRELATION PLOT, BUT THIS TIME WE TRY TO SEE THE 
CORRELATION BETWEEN VICTIM'S RACE AND GENDER THROUGH MOSAIC PLOT.

```{r}
plot4 <-mosaicplot(table(data$VIC_SEX,data$VIC_RACE),
           main="Mosaic for Race and Gender Correlation",
           xlab="Victim Gender",
           ylab="Victim Race",las = 1)
```

PLOT 5

IN PLOT 5 WE VISUALIZE SOME BOXPLOTS IN THE VICTIM'S RACE AND THEIR PRECINCT.

```{r}
data$PRECINCT<- as.numeric(data$PRECINCT)
plot5 <-data %>%
  ggplot(aes(x=PRECINCT,y=BORO))+
  geom_boxplot(,color="black",fill="grey")+
  labs(title = 'Boxplots of Precinct in every Neighborhood')+
  scale_x_continuous(breaks=seq(0,125,5))
plot5
```

PLOT 6

IN PLOT 6 WE HAVE SOME LINES THAT INDICATES TOTAL VICTIMS AND DISTINGUISH THEM 
PER THEIR AGE GROUP. OBVIOUSLY MOST OF THE VICTIMS ARE BETWEEN 18-45. HOWEVER,
WEE SEE THAT THERE ARE MANY TEENAGER VICTIMS.

```{r}
plot6 <-data %>%
  group_by(VIC_AGE_GROUP,PRECINCT)%>%
  summarize(count = n()) %>%
  ggplot(aes(x=PRECINCT, y=count,color=VIC_AGE_GROUP))+
  geom_line()+
  ggtitle("Total Victims per Age Group")+
  labs(x = "PRECINCT", y = "Total Victims")+
  scale_y_continuous(breaks=seq(0,650,50))+
  scale_x_continuous(breaks=seq(0,125,10))
plot6
```

PLOT 7 

WE HAVE AN INTERESTING PLOT IN PLOT 7, THAT INDICATE'S WHERE THE VICTIM WAS, 
WHEN IT WAS SHOT. WE DISTINGUISH THE VICTIMS THROUGH THEIR GENDER IN ORDER TO 
SEE WHERE MEN AND WOMEN USUALLY HANG OUT IN NEW YORK.

```{r}
plot7a <- data %>%
  select(LOCATION_DESC,VIC_SEX) %>%
  filter(LOCATION_DESC!="" & VIC_SEX=="M") %>%
  group_by(LOCATION_DESC) %>%
  summarise(Total_Men = n()) %>%
  arrange(desc(Total_Men)) %>%
  slice(1:8) %>%
  ggplot(aes(x = "", y = Total_Men, fill = LOCATION_DESC)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_brewer(palette = "Dark2") +
  ggtitle("Victim Men Location")

plot7b<- data %>%
  select(LOCATION_DESC,VIC_SEX) %>%
  filter(LOCATION_DESC!="" & VIC_SEX=="F") %>%
  group_by(LOCATION_DESC) %>%
  summarise(Total_Men = n()) %>%
  arrange(desc(Total_Men)) %>%
  slice(1:8) %>%
  ggplot(aes(x = "", y = Total_Men, fill = LOCATION_DESC)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_brewer(palette = "Dark2") +
  ggtitle("Victim Women Location")

grid.arrange(plot7a, plot7b, nrow = 2)
```

PLOT 8

IN THE FINAL PLOT, WE HAVE A LEAFLET THAT CLUSTER THE LOCATIONS OF THE SHOOTINGS,
AND WE SEE IN WHICH AREAS OF NEW YORK WE HAVE THE MOST OF THEM.

```{r}
# Transform our df in a sf object
data$Longitude<-as.numeric(data$Longitude)
data$Latitude<-as.numeric(data$Latitude)
sf_data <- st_as_sf(data,coords = c("Longitude","Latitude"),crs=4326)

# create leaflet map centered on New York
ny_map <- leaflet() %>% 
setView(lng = -74.0060, lat = 40.7128, zoom = 10)

# add tiles to the map
ny_map %>% 
  addTiles() %>% 
  addCircleMarkers(data = sf_data,
    radius = 2,
    label = ~ htmlEscape(VIC_RACE),
    clusterOptions = markerClusterOptions()
  ) 
```

